# Getting Started

Here is an overview of the process to automate flow cytometry data using R's `openCyto` and what you will need to successfully automate your own flow cytometry analysis. The general steps to accomplish this are as follows:

1. Read in a manually gated flowJo workspace in .wsp file format.
2. Parse raw FCS files from the read in workspace.
3. Visualize the manual gating template and resulting gates to verify gating scheme.
4. Create and read in a .csv gating template.
5. Automate gating.
6. Visualize automated gating template and gates to verify gating scheme. 
7. Extract population statistics and relevant information. 

This process is completed primarily with the `openCyto` package but calls upon other packages within the Bioconductor `openCyto` framework. Packages needed to complete this tutorial are listed at the end of this chapter. Descriptions of each function and R object used for this analysis are below. 

## Function and R Object Definitions
```{r var def table, echo = FALSE}
library(knitr)
vars <- data.frame(var_name = c("wsfile", "openWorkspace()", "ws",
                                "parseWorkspace()", "gating_set", "clone()",
                                "gh", "gt", "templateGen()", "gatingTemplate()",
                                "gating()", "plot()", "plotGate()"),
                    definition = c("flowJo .wsp file location", 
                                   "function used to read in **wsfile**",
                                   "read in data from flowJo",
                                   "function to extract FCS files from **ws**",
                                   "parsed FCS files to be gated",
                                   "function used to create a clone of **gating_set**",
                                   "subset of gating_set",
                                   ".csv gating template",
                                   "function used to generate a .csv template from existing manual gates",
                                   "function used to read in .csv template",
                                   "function used to apply gates to a gating set",
                                   "function to visualize gating tree",
                                   "function to visualize gates"))
kable(vars, align = c("c", "c", "c", "c", "c"),
      col.names = c("Function/Object Name", "Definition"))
```

## Required Packages and Installation  

Before getting started, install and load the following libraries into a new R script. As you will see below, this tutorial uses the development version of `openCyto`. Use the following to ensure the correct packages are installed.  

RStudio may also prompt you to download [XQuartz](https://www.xquartz.org/) and [XCode](https://developer.apple.com/xcode/), so it may be a good idea to go ahead and also download both before beginning.  

*To install*  
```{r install, warning=FALSE, eval=FALSE, message=FALSE, echo = TRUE}
devtools::install_github("RGLab/openCyto", ref = "trunk")
install.packages("data.table")
install.packages("flowWorkspace")
install.packages("flowCore")
install.packages("flowStats")
install.packages("flowClust")
install.packages("plyr")
```

*To load*  
```{r packages1, warning=FALSE, eval=FALSE, message=FALSE, echo = TRUE}
library(openCyto)
library(flowWorkspace)
library(data.table)
library(flowCore)
library(flowStats)
library(flowClust)
library(plyr)
```
