# Getting Started

Here is an overview of the process to automate flow cytometry data using R's `openCyto` and what you will need to successfully automate your own flow cytometry analysis. The general steps to accomplish this are as follows:

1. Read in a manually gated flowJo workspace in .wsp file format.
2. Parse raw FCS files from the read in workspace.
3. Visualize the manual gating template and resulting gates to verify gating scheme.
4. Create and read in a .csv gating template.
5. Automate gating.
6. Visualize automated gating template and gates to verify gating scheme. 
7. Extract population statistics and relevant information. 

This process is completed primarily with the `openCyto` package but calls upon other packages within the [Bioconductor](http://bioconductor.org/packages/release/bioc/html/openCyto.html) `openCyto` framework. Packages needed to complete this tutorial are listed at the end of this chapter. Descriptions of each function and R object used for this analysis are below. 


## Required Packages and Installation  

### Package descriptions  

Below is a description of each package used in this analysis. Code to install and use these packages will follow. Package descriptions taken from Bioconductor and CRAN.  

```{r  pack_def, echo = FALSE}
library(knitr)
vars <- data.frame(pack_name = c("openCyto", "flowWorkspace", "flowCore",
                                "flowStats", "flowClust", "data.table"),
                    definition = c("This package is designed to facilitate the automated gating methods in sequential way to mimic the manual gating strategy.", 
                                   "This package allows you to import basic flowJo workspaces into BioConductor and replicate the gating from flowJo using the flowCore functionality. Gating hierarchies, groups of samples, compensation, and transformation are performed so that the output matches the flowJo analysis.",
                                   "Provides S4 data structures and basic functions to deal with flow cytometry data.",
                                   "Methods and functionality to analyse flow data that is beyond the basic infrastructure provided by the flowCore package.",
                                   "Robust model-based clustering using a t-mixture model with Box-Cox transformation. ",
                                   "Fast aggregation of large data (e.g. 100GB in RAM), fast ordered joins, fast add/modify/delete of columns by group using no copies at all, list columns, friendly and fast character-separated-value read/write. Offers a natural and flexible syntax, for faster development."))
kable(vars, align = c("c", "c", "c", "c", "c"),
      col.names = c("Package Name", "Use"))
```

### Installation  

Install the following libraries into a new R script. As you will see below, this tutorial uses the development version of `openCyto`. It is important to use the development version of `openCyto` to remain up to date on any changes made by the developers of `openCyto`. Use the following to ensure the correct packages are installed. Installation will only need to be done once.  

*To install*  
```{r install, warning=FALSE, eval=FALSE, message=FALSE, echo = TRUE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install() 

BiocManager::install(c("openCyto", "flowWorkspace", "flowCore",
                       "flowStats", "flowClust"))

install.packages("data.table")
devtools::install_github("RGLab/openCyto", ref = "trunk")
```

RStudio may also prompt you to download [XQuartz](https://www.xquartz.org/) and [XCode](https://developer.apple.com/xcode/) based on your computer type, so it may be a good idea to go ahead and also download both.  

### Load packages  

Although installation only needs to be done once, packages will need to be reloaded each time you open an R session. At the beginning of each session, run the following code.  

*To load*  
```{r packages1, warning=FALSE, eval=FALSE, message=FALSE, echo = TRUE}
library(openCyto)
library(flowWorkspace)
library(data.table)
library(flowCore)
library(flowStats)
library(flowClust)
```
