# Working with your Manual Gating Scheme

Current methods for manually gating flow cytometry are both time consuming and costly, making automated gating an appealing option. The `openCyto` package allows users to take manually gated data from flowJo, reproduce those gates in R, and eventually automate the gating process. 

The first step in this process is to bring a pre-existing flowJo file into R in order to recreate the gating environment. The remainder of this chapter will detail the following:

1. Read in flowJo .wsp file  
2. Parse FCS files  
3. Visualize and verify manual gates 
4. Clone data for later automation

## Read in flowJo file

Within flowJo, tranformation, compensation, and gating can be saved as either .xml or .wsp filetypes. This tutorial will only detail steps from a *.wsp* filetype saved from flowJo. Note that many other tutorials begin from a *.xml* file, which seems to cause many errors during this analysis. Saving analysis within flowJo is detailed [here](http://docs.flowjo.com/vx/workspaces-and-samples/ws-savinganalysis/). 

The result of step 1 will be replication of the manual transformation, compensation, and gating from the flowJo workspace saved as an R object. Before you begin, be sure you have loaded the required packages outlined in the previous chapter.

Once all packages are loaded, save the .wsp file path as an R object called **wsfile**. Next, use `openWorkspace()` with your R object name to open the .wsp file in R, save this as **ws**. Here is an example of saving and opening **wsfile**. Following this step, **ws** will be saved as a flowWorkspace object containing groups of samples. 

```{r packages, warning=FALSE, message=FALSE, echo = TRUE}
library(openCyto)
library(flowWorkspace)
library(data.table)
library(flowCore)
library(flowStats)
library(flowClust)
library(plyr)
```
```{r manual gating data, warning=FALSE, message=FALSE}
wsfile <- "/Users/monhait/Desktop/group1_v_group2.wsp"
```
```{r open, warning=FALSE, message=FALSE}
ws <- openWorkspace(wsfile)
```

```{r look at ws, warning=FALSE, message=FALSE}
print(ws)
```

## Parse FCS files 

Once this file exists as an R object, the raw FCS files are then read using the `parseWorkspace` function. This function will read the FCS files and transform, compensate, and gate according to parameters defined from the .wsp flowJo workspace. The `parseWorskpace` call requires **ws** (the .wsp workspace that was just read in from flowJo) and the name of the samples to read in. Other options may be customized based on particular needs. A new R object named **gating_set** is then created and will be a GatingSet object. The **isNcdf = TRUE** call saves this output to disk rather that into memory because the files are large. Here is an example of parsing FCS files. After this, `attributes()` is used to examine the data.    

```{r parse_FCS, warning=FALSE, message=FALSE}
gating_set <- parseWorkspace(ws, name = "Samples", path = "/Users/monhait/Desktop/data/group1_v_group2", isNcdf = TRUE, sampNloc = 'sampleNode')
```
```{r rename, echo = FALSE}
newNames <- c("X_group1_1", "X_group1_2", "X_group1_3", "X_group1_4", "X_group1_5",
              "X_group2_1", "X_group2_2", "X_group2_3", "X_group2_4", "X_group2_5")
sampleNames(gating_set) <- newNames
pData(gating_set)[["name"]] <- newNames
```
```{r attributes, results = 'hide'}
attributes(gating_set)
```

## Visualize and Verify

It is helpful to now visualize both the gating template and actual gates on a subset of the data in order to verify the gating scheme. This will ensure consistency between the flowJo workspace and the manual gates recreated in R. First, save a subset of the gating_set as follows. The following saves the first FCS file of **gating_set** as **gh**. 

```{r save_subset, warning=FALSE, message=FALSE}
gh <- gating_set[[1]]
print(gh)
```

### plot()

The plot() function will visualize the current gating hierarchy. This can be done for the entire gating hierarchy or a specific population as seen below. 

```{r manual_hierarchy, warning=FALSE, message=FALSE}
plot(gh)
```

### plotGate()

The plotGate() function will gate the designated subset of your data according to parameters replicated from flowJo. 

```{r visualize_subset,warning=FALSE, message=FALSE}
flowWorkspace.par.set("plotGate", list(xlim = "data",
                                       ylim = "data"))
plotGate(gh)
```

**Note the use of `flowWorkspace.par.set()` here. Chapter 5 of this tutorial will discuss customizations such as this one.   

## Clone

Now, the manual gating scheme has been replicated in R using `openCyto` and verified for consistency with the original flowJo workspace. The final step in this chapter will clone **gating_set** into a new R object named **auto_gating**. 

```{r clone_data, warning=FALSE, message=FALSE}
auto_gating <- clone(gating_set)
print(auto_gating)
```



