# Automate Gating

## Load .csv into R  

As noted in the previous chapter, there is a sample gating template titled *"partial.csv"* with the sample data. This may serve as a guide to creating your own. When the .csv gating template is complete, it is then read into R and saved as `gt`, a gatingTemplate object. 

```{r read_csv, warning=FALSE, message=FALSE}
gt <- gatingTemplate("../tutorial/partial_gt.csv")
```

## Read in raw FCS files  

Now that the GatingTemplate object has been loaded into R, you will need to also load raw FCS files to perform the automated gating on. For gating, these files must be in a GatingSet object type. The path is saved as a character matrix of file names using `list.files()`. Here, this matrix is saved as `fcs_files`. Next, `read.ncdfFlowSet()` will save FCS files as a ncdfFlowSet object, here named `ncfs`. The `GatingSet` function will then save the FCS files as a GatingSet object, here named `gs_auto`. In this form, the FCS files can be input and gated. Use the following code to perform these steps.      

```{r raw_fcs, message=FALSE, warning=FALSE}
fcs_files <- list.files(path = "../tutorial/group1_v_group2", full.names = TRUE)
ncfs  <- read.ncdfFlowSet(files = fcs_files)
gs_auto <- GatingSet(ncfs)
```
```{r rename_new, echo = FALSE}
newNames <- c("X_group1_1", "X_group1_2", "X_group1_3", "X_group1_4", "X_group1_5",
              "X_group2_1", "X_group2_2", "X_group2_3", "X_group2_4", "X_group2_5")
sampleNames(gs_auto) <- newNames
pData(gs_auto)[["name"]] <- newNames
```
## Apply Gating  

At this point, you now have GatingTemplate and GatingSet object to be used for gating. Apply your GatingTemplate object to the GatingSet object, where x = GatingTemplate object and y = GatingSet object. Here, the GatingTemplate object is named `gt` and the GatingSet object is named `gs_auto`.     

```{r apply_gt, message=FALSE, warning=FALSE}
gating(x = gt, y = gs_auto)
```

    
## Plot Automated Gating  

Just as before, plot both the gating hierarchy and the automated gates. 

```{r plot_auto_gates, warning = FALSE}
plotGate(gs_auto[[1]])
```
```{r plot_auto_hierarchy, warning = FALSE}
plot(gs_auto[[1]])
```

## Population Statistics  

Both counts and frequencies of cell types in samples can be generated for analysis. Statistics can be generated from R or pulled directly from flowJo. To pull from flowJo, add `flowJo=TRUE` to either code chunk below.  

**Counts**  

```{r stats_count, warning=FALSE, message=FALSE}
head(getPopStats(gs_auto,statistic="count"))
```

**Frequencies**  

```{r stats_freq, warning=FALSE, message=FALSE}
head(getPopStats(gs_auto,statistic="freq"))
```


